Dispersion of short waves riding on longer waves

### 2024-05-10 paper/paper.tex
Add references

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
$k$ is the wavenumber, and $U$ is the mean advective current in the direction
of the wave propagation.
A current in the direction of the waves thus increases its absolution frequency,
wavenumber remaining unchanged.
An important limitation here is that $U$ must be slow-varying on the scales of
the wave period.
In the context of short waves riding on long waves, the mean advective current
is the horizontal near-surface orbital velocity of the long wave.

@@

As the short waves ride on the surface of the long wave,
they move in an accelerated reference frame and will thus experience effective
gravitational acceleration that changes in both space and time.
Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:

@@

\subsection{Wavenumber and amplitude modulation}

@@

\begin{figure}[h]
\label{fig:wavenumber_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_by_phase.png}
\caption{
  Comparison of relative short-wave modulation as function of the
  long-wave phase, for a moderately steep long wave of $a_L k_L = 0.1$.
  The dotted line shows the 1st-order approximated solution by L-HS,
  the dashed line shows the exact solution by L-HS,
  and the solid line shows the extended solution introduced in this paper.
}
\end{figure}

\subsection{Apparent gravity}

Long waves also induce orbital acceleration under their surface, so the short
waves experience an apparent gravity:

\begin{equation}
\label{eq:apparent_gravity1}
\widetilde{g} = g + \dfrac{\partial w_L}{\partial t}
\end{equation}
The positive sign in front of $\partial w_L/\partial t$ comes from our
coordinate-system convention where $w$ is defined positive-upward and $g$ is
defined positive-downward, and the fact that $\widetilde{g}$ is the apparent
gravity in the reference frame of the short waves.
Taking the long-wave vertical acceleration at the surface as:

\begin{equation}
\label{eq:w_L}
w_L = - a_L \omega_L e^{a_L k_L \cos{\psi}} \sin{\psi}
\end{equation}
we arrive at:

\begin{equation}
\label{eq:apparent_gravity2}
\widetilde{g}(\psi) = g \left[
  1 - a_L k_L e^{a_L k_L \cos{\psi}} \left(
    \cos{\psi} + a_L k_L \sin^2{\psi}
  \right)
\right]
\end{equation}
The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
so the apparent gravity of the short waves is increased at the long-wave crests
and decreased in the troughs.
The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
$\pi/2$ relative to the crests.
The second-order term thus increases the apparent gravity on the front and rear
faces of the long wave (Fig. \ref{fig:apparent_gravity}).

\begin{figure}[h]
\label{fig:apparent_gravity}
\centering
\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_phase.png}
\caption{
  Relative change in gravity of short waves due to a longer wave with steepness $a_L k_L = 0.2$,
  as a function of the long-wave phase and evaluated to the first (dashed line) and second (solid line) order in $a_L k_L$.
}
\end{figure}

\begin{figure}[h]
\label{fig:gravity_modulation_vs_ak}
\centering
\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_by_ak.png}
\caption{
  TODO
}
\end{figure}

>>>>

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
$k$ is the wavenumber, and $U$ is the mean advective current in the direction
of the wave propagation.
A current in the direction of the waves thus increases its absolution frequency,
wavenumber remaining unchanged.
An important limitation here is that $U$ must be slow-varying on the scales of
the wave period \cite{bretherton1968wavetrains}.
In the context of short waves riding on long waves, the mean advective current
is the horizontal near-surface orbital velocity of the long wave.

@@

As the short waves ride on the surface of the long wave,
they move in an accelerated reference frame and will thus experience effective
gravitational acceleration that changes in both space and time \cite{longuet1986eulerian,longuet1987propagation}.
Inserting Eq. (\ref{eq:dispersion}) into (\ref{eq:wave_action}) we obtain:

@@

\subsection{Wavenumber and amplitude modulation}
\label{subsection:analytical_solutions}

@@

\begin{figure}[h]
\label{fig:wavenumber_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_wavenumber_modulation_2panel.png}
\caption{
  Comparison of relative short-wave modulation as function of the
  long-wave phase, for a moderately steep long wave of $\varepsilon = 0.1$.
  The dotted line shows the 1st-order approximated solution by L-HS,
  the dashed line shows the exact solution by L-HS,
  and the solid line shows the extended solution introduced in this paper.
}
\end{figure}

\subsection{Gravity modulation}
\label{subsection:gravity_modulation}

Long waves induce vertical orbital accelerations, so the short
waves that ride on their surface move in an accelerated reference frame and
experience effective gravitational acceleration that is lower at the crests and
higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
A linearized (sine-like) wave has a modulation in gravity acceleration at its
surface:

\begin{equation}
\label{eq:effective_gravity}
\widetilde{g}(\psi) = g \left[
  1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
    \cos{\psi} + \varepsilon_L \sin^2{\psi}
  \right)
\right]
\end{equation}
The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
so the apparent gravity of the short waves is increased at the long-wave crests
and decreased in the troughs.
The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
$\pi/2$ relative to the crests.
The second-order term thus increases the apparent gravity on the front and rear
faces of the long wave (Fig. \ref{fig:apparent_gravity}).

\begin{figure}[h]
\label{fig:gravity_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_gravity_modulation_2panel.png}
\caption{
  Relative change in gravity of short waves due to a longer wave with steepness $a_L k_L = 0.2$,
  as a function of the long-wave phase and evaluated to the first (dashed line) and second (solid line) order in $a_L k_L$.
}
\end{figure}
